---
title: "Social network analysis in eSports Studies"
author: "Ilya Musabirov, Denis Bulygin, Viktor Karepin, Vadim Voskresenskii, Alexander Sirotkin"
output: html_document
---

# Installation of required R packages

First of all, we check which of the required packages are already installed.
Install missing packages, if any.

```{r}
list.of.packages <- c("magrittr", "dplyr", "igraph", "ggplot2", "ggraph")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) {install.packages(new.packages)}
rm(list = c("new.packages", "list.of.packages"))

library(magrittr)
library(dplyr)
library(igraph)
library(ggplot2)
library(ggraph)
```

# Short introduction to SNA packages (igraph, ggraph) for R 

Detailed introduction to SNA in R:
http://kateto.net/network-visualization
Page of ggraph, which is an extension for ggplot2 package:
https://github.com/thomasp85/ggraph

The objects we study in this tutorials are graphs (or networks). They consist of
a set of nodes and a set of edges. 
In the following command, we are assigning to the variable g a graph that has nodes
V = {1, 2, 3, 4, 5} and has edges E = {(1, 2),(1, 3),(2, 3),(3, 5)}.
The commands V(g) and E(g) print the list of nodes and edges of the graph g. Moreover, with the help of these commands you can assign attributes to nodes and edges.

```{r}
g = graph( c(1,2, 1,3, 2,3, 3,5))

V(g)
E(g)

V(g)$color = c("green", "green", "green", "green", "red")
E(g)$weight = c(1, 2, 4, 1)
```

The easiest way to draw a network is to use standart igraph plot function. The plot command has many parameters that control the behavior of the visualization.

```{r}
plot(g)
plot(g, vertex.color = V(g)$color, edge.width = E(g)$weight)
```

Moreover, you can use ggraph. It extends the grammar of graphics provided by ggplot2 to cover network data. ggraph is a flexible and powerful tool for network viualization and comes with a support for igraph data objects.

It works with the same three core concepts that are quite easy to understand:

- Layout - defines how nodes are placed on the plot.
- Nodes - the connected entities in the relational structure. These can be plotted using the geom_node_*() family of geoms. 
- Edges - the connections between the entities in the relational structure. These can be visualized using the geom_edge_*() family of geoms.

Let's create graph of highschool friendships and draw it using ggraph.

```{r}
graph <- graph_from_data_frame(highschool)
V(graph)$Popularity <- degree(graph, mode = 'in')

ggraph(graph, layout = 'kk') + 
    geom_edge_fan() + 
    geom_node_point(aes(size = Popularity)) + 
    facet_edges(~year)
```

# Read dataset 

In that tutorial we will use a dataset about co-usage of characters during Dota2 matches.
Dataset covers two last major competitions, which happened in Boston and Kiev.

Given csv-file (teams.csv) consist of edgelist with hero-hero pairs, which were used together in a match and metadata.
Meta:
- team - title of team, which picked that pair of heroes;
- competition - tournament, where the match took place;
- n_matches - number of matches, when that particular pair of heroes were picked together;
- wins - number of successful matches, when that pair of heroes were picked together;
- percentage_wins - percentage of successful matches, when that pair of heroes were picked together;

```{r}
edg = read.csv("./teams.csv", stringsAsFactors = FALSE)
```

# Comparison of hero picking stratagies of two finalists teams of Kiev Major Tournament

Let's take history of hero picking for only two teams ("OG Dota2" and "Virtus.pro") for competition held in Kiev. And load edgelist to igraph object.

```{r}
edg_filtered = edg %>%
  filter(team %in% c("OG Dota2", "Virtus.pro"),
         competition == "Kiev")

net = graph_from_data_frame(edg_filtered, directed = FALSE)
```

Next, we create a ggraph object with all hero-hero pairs for both teams. 
Transparency of edges is set to percentage of successful matches.
Width of edges corresponds to the number of matches played with that pair of heroes

```{r}
net_vis = ggraph(net, layout = 'fr') + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), colour = "steelblue3") + 
    geom_node_point(fill = "gray75", shape = 21, color = "gray30", size = 5) 

net_vis
```

Add text labels with heroes' titles.

```{r}
net_vis = net_vis + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE)

net_vis
```

Split network visualization by team to see the differences.

```{r}
net_vis = net_vis + 
  facet_edges(~ team)
    
net_vis
```

Add minor cosmetic changes.
We change the transparency scale and set background color and borders.

```{r}
net_vis = net_vis + 
  scale_edge_alpha(range = c(0.3, 0.9)) +
  theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis
```

# Comparison of hero picking stratagies of two finalists teams of Kiev Major Tournament. Before and after game update (patch).

Firstly, we take Virtus.pro team and load filtered edgelist to igraph.

```{r}
edg_filtered = edg %>%
  filter(team == "Virtus.pro")

net_virtus = graph_from_data_frame(edg_filtered, directed = FALSE)
```

Then, we pre-compute position of nodes on 2-dimensional space.
As we create separate visualizations, we have to freeze nodes' position in order to make the graphs comparable.

```{r}
net_virtus_layout = layout.fruchterman.reingold(net_virtus)
net_virtus_layout = as.data.frame(net_virtus_layout)
names(net_virtus_layout) = c("x", "y")
```

Draw network of Virtus.pro hero picking strategy for two tournaments

```{r}
net_vis_virtus = ggraph(net_virtus, layout = 'manual', node.position = net_virtus_layout) + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), show.legend = TRUE, colour = "steelblue3") + 
    geom_node_point(fill = "gray75", shape = 21, color = "gray30", size = 5) + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE) +
    scale_edge_alpha(range = c(0.3, 0.9)) +
    facet_edges(~competition) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis_virtus
```

Now, we take Virtus.pro team history only for Boston Major.

```{r}
net_virtus_b = delete.edges(net_virtus, which(E(net_virtus)$competition != "Boston"))
```

Next, we compute community detection to find densely connected subgraphs 
(combinations of heroes, which were more frequently taken together).
And remove clusters, consist of only one node.

```{r}
V(net_virtus_b)$clust = cluster_walktrap(net_virtus_b)$membership

V(net_virtus_b)$clust = ifelse(
  V(net_virtus_b)$clust %in% names(which(table(V(net_virtus_b)$clust) > 1)),
  V(net_virtus_b)$clust, NA)
```

Draw network of Virtus.pro hero picking strategy for Boston tournament (before Patch).

```{r}
net_vis_virtus_boston = ggraph(net_virtus_b, layout = 'manual', node.position = net_virtus_layout) + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), show.legend = TRUE, colour = "steelblue3") + 
    geom_node_point(aes(fill = as.factor(clust)), shape = 21, color = "gray30", size = 5) + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE) +
    scale_edge_alpha(range = c(0.3, 0.9)) +
    facet_edges(~competition) +
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis_virtus_boston
```

Now we take team history of Virtus.pro for Kiev Major (after Patch) and compute clusters again.

```{r}
net_virtus_k = delete.edges(net_virtus, which(E(net_virtus)$competition != "Kiev"))

V(net_virtus_k)$clust = cluster_walktrap(net_virtus_k)$membership
V(net_virtus_k)$clust = ifelse(
  V(net_virtus_k)$clust %in% names(which(table(V(net_virtus_k)$clust) > 1)),
  V(net_virtus_k)$clust, NA)
```

Finally, draw network of Virtus.pro hero picking strategy for Kiev tournaments. Now we can compare patterns of character picking for two periods. 

```{r}
net_vis_virtus_kiev = ggraph(net_virtus_k, layout = 'manual', node.position = net_virtus_layout) + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), show.legend = TRUE, colour = "steelblue3") + 
    geom_node_point(aes(fill = as.factor(clust)), shape = 21, color = "gray30", size = 5) + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE) +
    scale_edge_alpha(range = c(0.3, 0.9)) +
    facet_edges(~competition) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis_virtus_kiev
```

In addition, we can find a subgraph in which every vertex has at least degree k. 
And compare whether they changed after game update. 

```{r}
coreness(net_virtus_k) %>% 
  sort() %>% 
  tail(6)

coreness(net_virtus_b) %>% 
  sort() %>% 
  tail(6)
```


Do the same for OG Dota2 team

```{r}
edg_filtered = edg %>%
  filter(team == "OG Dota2")

net_og = graph_from_data_frame(edg_filtered, directed = FALSE)
```

Pre-compute layout.

```{r}
net_og_layout = layout.fruchterman.reingold(net_og)
net_og_layout = as.data.frame(net_og_layout)
names(net_og_layout) = c("x", "y")
```

Draw graph for both events.

```{r}
net_vis_og = ggraph(net_og, layout = 'manual', node.position = net_og_layout) + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), show.legend = TRUE, colour = "steelblue3") + 
    geom_node_point(fill = "gray75", shape = 21, color = "gray30", size = 5) + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE) +
    scale_edge_alpha(range = c(0.3, 0.9)) +
    facet_edges(~competition) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis_og
```

Take matches in Boston for OG team.

```{r}
net_og_b = delete.edges(net_og, which(E(net_og)$competition != "Boston"))
```

Community detection.

```{r}
V(net_og_b)$clust = cluster_walktrap(net_og_b)$membership
V(net_og_b)$clust = ifelse(
  V(net_og_b)$clust %in% names(which(table(V(net_og_b)$clust) > 1)),
  V(net_og_b)$clust, NA)
```

Draw! :)

```{r}
net_vis_og_boston = ggraph(net_og_b, layout = 'manual', node.position = net_og_layout) + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), show.legend = TRUE, colour = "steelblue3") + 
    geom_node_point(aes(fill = as.factor(clust)), shape = 21, color = "gray30", size = 5) + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE) +
    scale_edge_alpha(range = c(0.3, 0.9)) +
    facet_edges(~competition) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis_og_boston
```

Take matches in Kiev for OG team.

```{r}
net_og_k = delete.edges(net_og, which(E(net_og)$competition != "Kiev"))
```

Community detection. 
```{r}
V(net_og_k)$clust = cluster_walktrap(net_og_k)$membership
V(net_og_k)$clust = ifelse(
  V(net_og_k)$clust %in% names(which(table(V(net_og_k)$clust) > 1)),
  V(net_og_k)$clust, NA)
```

Draw and compare.

```{r}
net_vis_og_kiev = ggraph(net_og_k, layout = 'manual', node.position = net_og_layout) + 
    geom_edge_fan(aes(alpha = percentage_wins, width = n_matches), show.legend = TRUE, colour = "steelblue3") + 
    geom_node_point(aes(fill = as.factor(clust)), shape = 21, color = "gray30", size = 5) + 
    geom_node_text(aes(label = name), color = 'black', size = 3, repel = TRUE) +
    scale_edge_alpha(range = c(0.3, 0.9)) +
    facet_edges(~competition) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

net_vis_og_kiev
```

# Things to do...
1. 
2. Take one more Dota2 team (e.g. "Invictus Gaming", "Natus Vincere", "Team NP" or "Team Secret"), analyse its' hero picking pattern and compare with the cases of "OG Dota2" and "Virtus.pro".

